
---
title: "HCC Survival Analysis"
author: "Ana Saeedi"
date: "`r Sys.Date()`"
output:
  pdf_document:
    latex_engine: xelatex
    toc: true
    toc_depth: 3
    number_sections: true
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    df_print: paged
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 8,
  fig.height = 6
)
```

# Overview

This R Markdown document restructures the original HCC survival analysis script into a report format so the code, tables, model output, and plots can all be shown together after knitting.

# Load required packages

```{r packages}
# install.packages("labelled")
# install.packages("gmodels")
# install.packages("survminer")
# install.packages("survRM2")

library(haven)    
library(dplyr)    
library(survival)
library(survminer)
library(labelled)  
library(readr)     
library(scales)
library(forcats)
library(gmodels)
library(stringr)
library(tibble)
library(lubridate)
```

# Import dataset

```{r import-data}
file_path <- "C:/Users/saeeda1/OneDrive - UMass Chan Medical School/Documents/independent study/cancer project/HCC_4_6_SASPrep.sas7bdat"

chart_data <- read_sas(file_path)
glimpse(chart_data)

hcc_data <- read_sas(file_path)
```

# Data preparation

## Rename selected variables

```{r rename-vars}
chart_data <- hcc_data %>%
  mutate(
    residual_recurrence = residual_reccurence,
    date_ldt1 = date_LTD1,
    n_hcc_tx_ldt1 = n_hcc_tx_LTD1
  ) %>%
  select(-residual_reccurence, -date_LTD1, -n_hcc_tx_LTD1)
```

## Apply variable labels

```{r variable-labels}
var_labels <- c(
  unique_id = "Unique ID",
  study_id = "Study ID",
  unique = "Unique Patient",
  age_first_SBRT = "Age first SBRT",
  Milan_criteria = "Inside Milan",
  pvtt = "PVTT",
  etiology_1 = "Etiology 1",
  etiology_2 = "Etiology 2",
  ascites = "Ascites",
  listed = "Listed",
  n_tumor_LDT1 = "No. Tumors LDT1",
  date_LTD1 = "Date LDT1",
  n_hcc_tx_LTD1 = "No. HCC tumor tx LDT1",
  type_LDT1 = "Type LDT1",
  size_tx_lesion = "Size Tx Lesion",
  date_LDT2 = "Date LDT2",
  type_LDT2 = "Type LDT2",
  date_LDT3 = "Date LDT3",
  type_LDT3 = "Type LDT3",
  date_LDT4 = "Date LDT4",
  type_LDT4 = "Type LDT4",
  prior_LDT = "No. prior LDTs",
  date_last_LDT = "Date last LDT",
  last_LDT = "Last LDT",
  Desc_Salv_Consol = "SBRT Description",
  n_hcc__time_sbrt = "No. HCC tumors at SBRT",
  n_hcc_tx_sbrt = "No. HCC tumors treated SBRT",
  hcc_size_sbrt = "Size of SBRT treated lesion",
  cp_prior_group = "CP Group",
  afp_prior_sbrt = "AFP Prior to SBRT",
  td = "TD",
  n_fx = "No. Fx",
  eqd2 = "EQD2",
  sbrt_start = "SBRT Start",
  sbrt_end = "SBRT End",
  gtv = "GTV Vol",
  itv = "ITV Vol",
  ptv = "PTV Vol",
  mld = "MLD",
  afp_1_mo_after = "AFP 1 mo after SBRT",
  cp_comp_base_6_mo = "CP comparision baseline-6 months",
  cp_miss_data = "CP comparision missing data",
  gd_3_tx = "Gr3+ tox",
  elpd = "ELPD",
  elpd_date = "Date ELPD",
  dm = "DM",
  dm_date = "Date DM",
  lc_censor = "LC Censor Date (after SBRT)",
  lc_time = "LC Time (days)",
  lc_censor_reason = "LC Censor Reason",
  resp = "Response to SBRT at censor",
  radio_resp = "Radiographic Respose",
  path_resp = "Path response",
  transplant = "Transplant",
  os_censor = "OS Censor Date",
  os_censor_reason = "OS Censor Reason",
  os_cmt_salv = "OS (CMT/salv (T0 date LDT1)",
  residual_reccurence = "Salvage only",
  cmt_diag_test = "CMT Diagnostic Test before SBRT"
)

chart_data <- set_variable_labels(hcc_data, !!!var_labels)
```

## Optional CSV export

```{r export-csv}
write_csv(chart_data, "C:/Users/saeeda1/OneDrive - UMass Chan Medical School/Documents/independent study/cancer project/HCC_4_6_SASPrep.csv")
```

## Recode factors

```{r recode-factors}
chart_data <- chart_data %>%
  mutate(
    Milan_criteria = factor(Milan_criteria, levels = c(0, 1), labels = c("No", "Yes")),
    pvtt = factor(pvtt, levels = c(0, 1), labels = c("No", "Yes")),
    ascites = factor(ascites, levels = c(0, 1), labels = c("No", "Yes")),
    listed = factor(listed, levels = c("0", "1"), labels = c("No", "Yes")),
    elpd = factor(elpd, levels = c(0, 1), labels = c("No", "Yes")),
    dm = factor(dm, levels = c(0, 1), labels = c("No", "Yes")),
    transplant = factor(transplant, levels = c(0, 1), labels = c("No", "Yes")),
    Desc_Salv_Consol = factor(Desc_Salv_Consol, levels = c(1, 2), labels = c("Salvage", "Consolidate/CMV")),
    cp_prior_group = factor(cp_prior_group, levels = c(1, 2), labels = c("A", "B/C")),
    cp_comp_base_6_mo = factor(cp_comp_base_6_mo, levels = c(0, 1, 2, 3),
                               labels = c("Missing", "Progression", "Stable", "Improve")),
    cp_miss_data = factor(cp_miss_data, levels = c(4, 5, 6),
                          labels = c("Deceased", "Transplant", "Short F/U")),
    gd_3_tx = factor(gd_3_tx, levels = c(0, 1, 3), labels = c("No", "Yes", "Unknown")),
    residual_reccurence = factor(residual_reccurence, levels = c(1, 2, 3),
                                 labels = c("Residual", "Recurrence", "Not included")),
    cmt_diag_test = factor(cmt_diag_test, levels = c(0, 1, 3),
                           labels = c("No", "Yes", "Not included")),
    etiology_1 = factor(etiology_1, levels = 1:9,
                        labels = c("HCV", "ETOH", "HBV", "IC", "NASH", "Hemo", "Autoimmune", "PBC", "Other")),
    etiology_2 = factor(etiology_2, levels = 1:9,
                        labels = c("HCV", "ETOH", "HBV", "IC", "NASH", "Hemo", "Autoimmune", "PBC", "Other"))
  )
```

## Create first-record-per-patient dataset

```{r patient-level}
studyid_wise <- chart_data %>%
  arrange(study_id) %>%
  group_by(study_id) %>%
  slice(1) %>%
  ungroup()
```

# Descriptive analyses

## Group counts

```{r group-counts}
CrossTable(studyid_wise$Desc_Salv_Consol, chisq = TRUE)
CrossTable(chart_data$Desc_Salv_Consol, chisq = TRUE)
```

## Age at first SBRT

```{r age-analysis}
hist(chart_data$age_first_SBRT, breaks = 10, main = "Age at First SBRT", xlab = "Age")

shapiro.test(chart_data$age_first_SBRT)
summary(chart_data$age_first_SBRT)
sd(chart_data$age_first_SBRT)

studyid_wise %>%
  group_by(Desc_Salv_Consol) %>%
  summarise(
    median_age = median(age_first_SBRT, na.rm = TRUE),
    Q1 = quantile(age_first_SBRT, 0.25, na.rm = TRUE),
    Q3 = quantile(age_first_SBRT, 0.75, na.rm = TRUE),
    .groups = "drop"
  )

t.test(age_first_SBRT ~ Desc_Salv_Consol, data = studyid_wise)
wilcox.test(age_first_SBRT ~ Desc_Salv_Consol, data = studyid_wise, exact = FALSE)
```

## Tumor size at SBRT

```{r tumor-size-analysis}
hist(chart_data$hcc_size_sbrt, breaks = 10, main = "Tumor size", xlab = "Size")

shapiro.test(chart_data$hcc_size_sbrt)

chart_data %>%
  group_by(Desc_Salv_Consol) %>%
  summarise(
    median_size = median(hcc_size_sbrt, na.rm = TRUE),
    Q1 = quantile(hcc_size_sbrt, 0.25, na.rm = TRUE),
    Q3 = quantile(hcc_size_sbrt, 0.75, na.rm = TRUE),
    .groups = "drop"
  )

t.test(hcc_size_sbrt ~ Desc_Salv_Consol, data = chart_data)
wilcox.test(hcc_size_sbrt ~ Desc_Salv_Consol, data = studyid_wise, exact = FALSE)
```

## Selected categorical comparisons

```{r categorical-comparisons}
table1 <- table(studyid_wise$etiology_1, studyid_wise$Desc_Salv_Consol)
chisq.test(table1)

table(studyid_wise$etiology_2, studyid_wise$Desc_Salv_Consol)

fisher.test(table(studyid_wise$pvtt, studyid_wise$Desc_Salv_Consol))
chisq.test(table(chart_data$Milan_criteria, chart_data$Desc_Salv_Consol))
chisq.test(table(studyid_wise$listed, studyid_wise$Desc_Salv_Consol))
chisq.test(table(studyid_wise$transplant, studyid_wise$Desc_Salv_Consol))
chisq.test(table(studyid_wise$cp_prior_group, studyid_wise$Desc_Salv_Consol))
fisher.test(table(studyid_wise$cp_comp_base_6_mo, studyid_wise$Desc_Salv_Consol))
fisher.test(table(studyid_wise$gd_3_tx, studyid_wise$Desc_Salv_Consol))
```

## Continuous summary table

```{r summary-table}
chart_data <- chart_data %>%
  mutate(gtv = as.numeric(gtv))

chart_data %>%
  summarise(
    median_gtv = median(gtv, na.rm = TRUE),
    Q1_gtv = quantile(gtv, 0.25, na.rm = TRUE),
    Q3_gtv = quantile(gtv, 0.75, na.rm = TRUE),
    median_ptv = median(ptv, na.rm = TRUE),
    median_td = median(td, na.rm = TRUE),
    median_n_fx = median(n_fx, na.rm = TRUE),
    median_eqd2 = median(eqd2, na.rm = TRUE),
    median_mld = median(mld, na.rm = TRUE)
  )

vars <- c("gtv", "ptv", "td", "n_fx", "eqd2", "mld")
results_list <- list()

for (var in vars) {
  values <- chart_data[[var]]
  group <- chart_data$Desc_Salv_Consol

  df <- data.frame(value = values, group = group) %>% 
    filter(!is.na(value), !is.na(group))

  summary_stats <- df %>%
    group_by(group) %>%
    summarise(
      Median = median(value),
      Q1 = quantile(value, 0.25),
      Q3 = quantile(value, 0.75),
      .groups = "drop"
    )

  p_val <- if (length(unique(df$group)) == 2) {
    wilcox.test(value ~ group, data = df)$p.value
  } else {
    kruskal.test(value ~ group, data = df)$p.value
  }

  results_list[[var]] <- data.frame(
    Variable = var,
    Group1 = summary_stats$group[1],
    Median1 = summary_stats$Median[1],
    IQR1 = paste0("(", summary_stats$Q1[1], "–", summary_stats$Q3[1], ")"),
    Group2 = summary_stats$group[2],
    Median2 = summary_stats$Median[2],
    IQR2 = paste0("(", summary_stats$Q1[2], "–", summary_stats$Q3[2], ")"),
    P_value = round(p_val, 3)
  )
}

summary_table <- do.call(rbind, results_list)
summary_table
```

## Total summary table

```{r total-summary-table}
vars <- c("gtv", "ptv", "td", "n_fx", "eqd2", "mld")
results_list <- list()

for (var in vars) {
  values <- chart_data[[var]]
  group <- chart_data$Desc_Salv_Consol

  df <- data.frame(value = values, group = group) %>% 
    filter(!is.na(value), !is.na(group))

  summary_stats <- df %>%
    group_by(group) %>%
    summarise(
      Median = median(value),
      Q1 = quantile(value, 0.25),
      Q3 = quantile(value, 0.75),
      .groups = "drop"
    )

  total_summary <- data.frame(
    group = "Total",
    Median = median(df$value),
    Q1 = quantile(df$value, 0.25),
    Q3 = quantile(df$value, 0.75)
  )

  p_val <- if (length(unique(df$group)) == 2) {
    wilcox.test(value ~ group, data = df)$p.value
  } else {
    kruskal.test(value ~ group, data = df)$p.value
  }

  combined <- bind_rows(summary_stats, total_summary) %>%
    mutate(
      Variable = var,
      IQR = paste0("(", round(Q1, 1), "–", round(Q3, 1), ")"),
      P_value = ifelse(group == "Total", NA, round(p_val, 3))
    ) %>%
    select(Variable, Group = group, Median, IQR, P_value)

  results_list[[var]] <- combined
}

final_summary <- bind_rows(results_list)
final_summary
```

## Correct SBRT end date for one observation

```{r fix-sbrt-end}
hcc_data <- hcc_data %>%
  mutate(
    sbrt_end = if_else(
      unique_id == 77,
      as.Date("2018-10-19"),
      sbrt_end
    )
  )

hcc_data %>%
  filter(unique_id == 77) %>%
  select(unique_id, date_last_LDT, sbrt_end)
```

## Additional descriptive tables

```{r table-6-7}
table(chart_data$prior_LDT)
fisher.test(table(chart_data$prior_LDT, chart_data$Desc_Salv_Consol))

chart_data %>%
  summarise(
    median_prior_LDT = median(prior_LDT, na.rm = TRUE)
  )

table(chart_data$last_LDT, chart_data$Desc_Salv_Consol)
fisher.test(table(chart_data$last_LDT, chart_data$Desc_Salv_Consol))
```

# Local control analysis

## Build analysis dataset

```{r local-control-dataset}
model_set1 <- chart_data %>%
  mutate(
    time_lc_yr = as.numeric(difftime(lc_censor, sbrt_end, units = "days")) / 365,
    resp = coalesce(resp, ""),
    resp_fail = if_else(str_to_upper(str_trim(resp)) == "CPD", 1, 0)
  )

table(model_set1$resp, model_set1$resp_fail, useNA = "ifany")
table(model_set1$resp, useNA = "ifany")
table(model_set1$resp_fail, useNA = "ifany")
```

## Truncated 3-year event indicator

```{r local-control-3yr}
model_set1_3yr <- model_set1 %>%
  mutate(
    time_3yr = pmin(time_lc_yr, 3, na.rm = TRUE),
    status_3yr = ifelse(resp_fail == 1 & time_lc_yr <= 3, 1, 0)
  )

table(model_set1_3yr$status_3yr, useNA = "ifany")
table(model_set1$resp_fail, model_set1$Desc_Salv_Consol, useNA = "ifany")
```

## Follow-up summary

```{r local-control-followup}
summary(model_set1$time_lc_yr)
max(model_set1$time_lc_yr, na.rm = TRUE)

model_set1 %>% 
  summarise(
    n          = n(),
    min_follow = min(time_lc_yr, na.rm = TRUE),
    q1         = quantile(time_lc_yr, .25, na.rm = TRUE),
    median     = median(time_lc_yr, na.rm = TRUE),
    q3         = quantile(time_lc_yr, .75, na.rm = TRUE),
    max_follow = max(time_lc_yr, na.rm = TRUE)
  )
```

## Overall local control KM

```{r lc-km-overall}
surv_obj <- Surv(time = model_set1$time_lc_yr,
                 event = model_set1$resp_fail)

fit_km <- survfit(surv_obj ~ 1, data = model_set1)

plot_km <- ggsurvplot(
  fit_km,
  conf.int      = FALSE,
  risk.table    = TRUE,
  xlab          = "Years from First Recurrence",
  ylab          = "Local Control (%)",
  title         = "Kaplan–Meier Curve for local control",
  xlim          = c(0, 4),
  break.time.by = 1,
  ggtheme       = theme_minimal()
)

plot_km$plot <- plot_km$plot +
  scale_x_continuous(limits = c(0, 4), breaks = 0:4) +
  theme(
    panel.border     = element_rect(color = "black", fill = NA, linewidth = 0.8),
    panel.background = element_blank(),
    panel.grid       = element_blank()
  )

print(plot_km)

summary(fit_km)
summary(fit_km, times = c(1, 2, 3, 4))
median_lc <- summary(fit_km)$table["median"]
median_lc
```

## Local control by treatment group

```{r lc-km-group}
surv_obj_full <- Surv(time = model_set1$time_lc_yr, event = model_set1$resp_fail)

fit_km_full <- survfit(surv_obj_full ~ 1, data = model_set1, conf.type = "log-log")
summary(fit_km_full, times = c(1, 2, 3, 4))

fit_stratified <- survfit(Surv(time_lc_yr, resp_fail) ~ Desc_Salv_Consol, data = model_set1)
summary(fit_stratified)

plot_km <- ggsurvplot(
  fit_stratified,
  conf.int   = FALSE,
  risk.table = TRUE,
  pval       = TRUE,
  xlab       = "Years from First Recurrence",
  ylab       = "Local Control (%)",
  title      = "Kaplan–Meier Plot (display truncated at 4 years)",
  xlim       = c(0, 4),
  break.time.by = 1,
  legend.title  = "Group",
  legend.labs   = c("Salvage", "CMT"),
  ggtheme       = theme_minimal()
)

plot_km$plot <- plot_km$plot +
  scale_x_continuous(limits = c(0, 4), breaks = 0:4) +
  theme(
    panel.border     = element_rect(color = "black", fill = NA, linewidth = 0.8),
    panel.background = element_blank(),
    panel.grid       = element_blank()
  )

print(plot_km)
```

## 4-year censored local control by group

```{r lc-km-group-4yr}
model_set1$time_lc_yr_4yr  <- pmin(model_set1$time_lc_yr, 4)
model_set1$event_4yr       <- ifelse(model_set1$time_lc_yr > 4, 0, model_set1$resp_fail)

fit_stratified <- survfit(Surv(time_lc_yr_4yr, event_4yr) ~ Desc_Salv_Consol, data = model_set1)

plot_km <- ggsurvplot(
  fit_stratified,
  conf.int      = FALSE,
  risk.table    = TRUE,
  pval          = TRUE,
  xlab          = "Years from First Recurrence",
  ylab          = "Local Control (%)",
  title         = "Kaplan–Meier Plot for overall Local Control by group",
  xlim          = c(0, 4),
  break.time.by = 1,
  legend.title  = "Group",
  legend.labs   = c("Salvage", "CMT"),
  ggtheme       = theme_minimal()
)

plot_km$plot <- plot_km$plot +
  scale_x_continuous(limits = c(0, 4), breaks = seq(0, 4, 1)) +
  theme(
    panel.border     = element_rect(color = "black", fill = NA, linewidth = 0.8),
    panel.background = element_blank(),
    panel.grid       = element_blank()
  )

print(plot_km)
```

## Local control probability table and RMST

```{r lc-prob-rmst}
fit_stratified <- survfit(Surv(time_lc_yr, resp_fail) ~ Desc_Salv_Consol, data = model_set1)

km_summary <- summary(
  fit_stratified,
  times = c(1, 2, 3, 4),
  censored = TRUE,
  extend = TRUE
)

fit_stratified_loglog <- survfit(
  surv_obj_full ~ Desc_Salv_Consol,
  data = model_set1,
  conf.type = "log-log"
)

km_summary_1 <- summary(fit_stratified_loglog, times = c(1, 2, 3, 4))
km_summary_1

km_df <- data.frame(
  Time_Year = km_summary$time,
  Group     = gsub("Desc_Salv_Consol=", "", km_summary$strata),
  N_Risk    = km_summary$n.risk,
  N_Event   = km_summary$n.event,
  Survival_Percent = round(100 * km_summary$surv, 1),
  CI_Lower_Percent = round(100 * km_summary$lower, 1),
  CI_Upper_Percent = round(100 * km_summary$upper, 1)
)

print(km_df)

tau_lc <- 2.5

sum_lc_overall <- summary(fit_km, rmean = tau_lc)$table

overall_lc_rmst <- tibble(
  group     = "Overall",
  n         = as.numeric(sum_lc_overall["n"]),
  tau       = tau_lc,
  rmst      = as.numeric(sum_lc_overall["rmean"]),
  se_rmst   = as.numeric(sum_lc_overall["se(rmean)"])
) |>
  mutate(
    lower_95 = rmst - 1.96 * se_rmst,
    upper_95 = rmst + 1.96 * se_rmst
  ) |>
  mutate(across(c(tau, rmst, se_rmst, lower_95, upper_95), ~ round(., 2)))

overall_lc_rmst

rmst_lc_tbl <- summary(fit_stratified, rmean = tau_lc)$table |>
  as.data.frame() |>
  rownames_to_column("group") |>
  transmute(
    group,
    n        = records,
    events   = events,
    tau      = tau_lc,
    rmst     = rmean,
    se_rmst  = `se(rmean)`,
    lower_95 = rmean - 1.96 * `se(rmean)`,
    upper_95 = rmean + 1.96 * `se(rmean)`
  ) |>
  mutate(across(c(tau, rmst, se_rmst, lower_95, upper_95), ~ round(., 2)))

rmst_lc_tbl
```

# Overall survival analysis

## Build OS dataset

```{r os-dataset}
model_overall <- studyid_wise %>%
  mutate(
    time_os = as.numeric(os_censor - date_LTD1) / 365,
    flg_over = ifelse(os_censor_reason == "died", 1, 0)
  )
table(model_overall$flg_over, useNA = "ifany")
nrow(model_overall)

model_overall <- studyid_wise %>%
  mutate(
    time_os   = as.numeric(os_censor - date_LTD1) / 365,
    status_os = ifelse(os_censor_reason == "died", 1, 0)
  )

table(model_overall$status_os, useNA = "ifany")
summary(model_overall$time_os)
```

## Overall OS KM, median, RMST

```{r os-km-overall}
fit_os <- survfit(Surv(time_os, status_os) ~ 1, data = model_overall)

plot_os <- ggsurvplot(
  fit_os,
  conf.int      = FALSE,
  risk.table    = TRUE,
  xlab          = "Years from Index Date",
  ylab          = "Overall Survival (%)",
  title         = "Kaplan–Meier Curve for Overall Survival",
  xlim          = c(0, 4),
  break.time.by = 1,
  ggtheme       = theme_minimal()
)

plot_os$plot <- plot_os$plot +
  scale_x_continuous(limits = c(0, 4), breaks = 0:4) +
  theme(
    panel.border     = element_rect(color = "black", fill = NA, linewidth = 0.8),
    panel.background = element_blank(),
    panel.grid       = element_blank()
  )

print(plot_os)

times <- c(1, 2, 3, 4)
km_at_t <- summary(fit_os, times = times)

prob_tbl <- data.frame(
  time      = km_at_t$time,
  surv_prob = km_at_t$surv,
  lower_95  = km_at_t$lower,
  upper_95  = km_at_t$upper
)

prob_tbl

med_vec <- summary(fit_os)$table[c("median", "0.95LCL", "0.95UCL")]

median_tbl <- data.frame(
  median_surv = med_vec["median"],
  lower_95    = med_vec["0.95LCL"],
  upper_95    = med_vec["0.95UCL"]
)

median_tbl

tau_os <- 7.0

sum_os_7 <- summary(fit_os, rmean = tau_os)$table
rmst_overall_os_7 <- data.frame(
  tau        = tau_os,
  rmst_total = as.numeric(sum_os_7["rmean"]),
  se_rmst    = as.numeric(sum_os_7["se(rmean)"])
)

rmst_overall_os_7$lower_95 <- with(rmst_overall_os_7, rmst_total - 1.96 * se_rmst)
rmst_overall_os_7$upper_95 <- with(rmst_overall_os_7, rmst_total + 1.96 * se_rmst)
rmst_overall_os_7[] <- lapply(rmst_overall_os_7, \(x) if(is.numeric(x)) round(x, 2) else x)

rmst_overall_os_7
```

## Reverse KM follow-up

```{r reverse-km-followup}
model_overall <- model_overall %>%
  mutate(followup_surv = Surv(time_os, status_os == 0))

rev_km_fit_total <- survfit(followup_surv ~ 1, data = model_overall, conf.type = "plain")
rev_km_total <- summary(rev_km_fit_total)$table[c("median", "0.95LCL", "0.95UCL")]

followup_total_tbl <- data.frame(
  Group            = "Total",
  Median_Followup  = rev_km_total["median"],
  Lower_95_CI      = rev_km_total["0.95LCL"],
  Upper_95_CI      = rev_km_total["0.95UCL"]
)

rev_km_fit_by_group <- survfit(followup_surv ~ Desc_Salv_Consol, data = model_overall, conf.type = "plain")
group_summaries <- summary(rev_km_fit_by_group)$table

followup_group_tbl <- data.frame(
  Group            = rownames(group_summaries),
  Median_Followup  = group_summaries[, "median"],
  Lower_95_CI      = group_summaries[, "0.95LCL"],
  Upper_95_CI      = group_summaries[, "0.95UCL"]
)

full_followup_tbl <- bind_rows(followup_total_tbl, followup_group_tbl)
full_followup_tbl

ggsurvplot(
  rev_km_fit_total,
  conf.int = TRUE,
  xlab = "Follow-up Time (Years)",
  ylab = "Proportion Remaining Unfollowed (Reverse KM)",
  title = "Reverse Kaplan–Meier Curve for Follow-Up Time"
)
```

## Overall survival by treatment group

```{r os-km-group}
fit_os <- survfit(Surv(time_os, status_os) ~ 1, data = model_overall)
model_overall$time_os_4yr <- pmin(model_overall$time_os, 4)
model_overall$event_os_4yr <- ifelse(model_overall$time_os > 4, 0, model_overall$status_os)

fit_stratified_os <- survfit(Surv(time_os_4yr, event_os_4yr) ~ Desc_Salv_Consol, data = model_overall)

plot_km <- ggsurvplot(
  fit_stratified_os,
  conf.int      = FALSE,
  risk.table    = TRUE,
  pval          = TRUE,
  xlab          = "Years from First Recurrence",
  ylab          = "Overal Survival (%)",
  title         = "Kaplan–Meier Plot for overall survival by group",
  xlim          = c(0, 4),
  break.time.by = 1,
  legend.title  = "Group",
  legend.labs   = c("Salvage", "CMT"),
  ggtheme       = theme_minimal()
)

plot_km$plot <- plot_km$plot +
  scale_x_continuous(limits = c(0, 4), breaks = seq(0, 4, 1)) +
  theme(
    panel.border     = element_rect(color = "black", fill = NA, linewidth = 0.8),
    panel.background = element_blank(),
    panel.grid       = element_blank()
  )

print(plot_km)
```



# Subgroup analysis

## Treatment interval subgroup for local control

```{r subgroup-local-control}
model_subgroup <- chart_data %>%
  mutate(
    months_treat_time = as.numeric(difftime(sbrt_end, date_last_LDT, units = "days")) / 30.44,
    subgroup_flg = case_when(
      months_treat_time <  6 ~ 1,
      months_treat_time >= 6 ~ 2,
      TRUE                   ~ NA_real_
    ),
    time_lc_yr = as.numeric(difftime(lc_censor, sbrt_end, units = "days")) / 365,
    resp_fail  = if_else(str_to_upper(str_trim(coalesce(resp, ""))) == "CPD", 1, 0),
    treatment_group = factor(
      subgroup_flg,
      levels = c(1, 2),
      labels = c("<6 months treatment", "≥6 months treatment")
    )
  )

surv_obj <- Surv(time = model_subgroup$time_lc_yr,
                 event = model_subgroup$resp_fail)

fit_km_sg <- survfit(surv_obj ~ treatment_group, data = model_subgroup)

summary(fit_km_sg, times = c(1, 2, 3, 4))
summary(fit_km_sg, times = 1:4, conf.type = "plain")
summary(fit_km_sg, conf.type = "plain")$table

max_by_arm <- with(model_subgroup,
                   tapply(time_lc_yr, treatment_group, max, na.rm = TRUE))
max_by_arm

tau <- 4.5
tbl <- summary(fit_km_sg, rmean = tau)$table

rmst_tbl <- tbl |>
  as.data.frame(check.names = FALSE) |>
  tibble::rownames_to_column("group") |>
  dplyr::transmute(
    group     = sub("^.*=", "", group),  
    tau       = tau,
    rmst      = rmean,
    se_rmst   = `se(rmean)`,
    lower_95  = rmst - 1.96 * se_rmst,
    upper_95  = rmst + 1.96 * se_rmst
  )

rmst_tbl
```

## Function to pick tau before heavy censoring

```{r tau-function}
cutoff_tau_before_heavy_censoring <- function(fit, by = 0.25, risk_threshold = 0.10) {
  s_all <- summary(fit)
  t_max <- min(tapply(s_all$time, s_all$strata, max, na.rm = TRUE), na.rm = TRUE)
  times <- seq(0, t_max, by = by)

  s <- summary(fit, times = times)

  df <- tibble(
    time     = s$time,
    strata   = s$strata,
    n_risk   = s$n.risk,
    n_event  = s$n.event,
    n_censor = s$n.censor
  )

  n0 <- df %>%
    group_by(strata) %>%
    summarise(N0 = max(n_risk, na.rm = TRUE), .groups = "drop")

  df2 <- df %>%
    inner_join(n0, by = "strata") %>%
    mutate(frac_at_risk = n_risk / N0)

  limiting <- df2 %>%
    group_by(time) %>%
    summarise(min_frac_at_risk = min(frac_at_risk, na.rm = TRUE), .groups = "drop")

  tau_suggested <- limiting %>%
    filter(min_frac_at_risk >= risk_threshold) %>%
    summarise(tau = max(time, na.rm = TRUE)) %>%
    pull(tau)

  list(
    tau_suggested       = tau_suggested,
    limiting_support    = limiting,
    per_time_support    = df2,
    t_max_shortest      = t_max,
    risk_threshold_used = risk_threshold,
    grid_by             = by
  )
}

out_sg <- cutoff_tau_before_heavy_censoring(fit_km_sg, by = 0.25, risk_threshold = 0.10)
out_sg$tau_suggested
head(out_sg$limiting_support, 10)

out_sg_20 <- cutoff_tau_before_heavy_censoring(fit_km_sg, by = 0.25, risk_threshold = 0.20)
out_sg_20$tau_suggested

tbl1 <- summary(fit_km_sg, rmean = out_sg$tau_suggested)$table

rmst_tbl <- tbl1 |>
  as.data.frame(check.names = FALSE) |>
  tibble::rownames_to_column("group") |>
  dplyr::transmute(
    group     = sub("^.*=", "", group),  
    tau       = tau,
    rmst      = rmean,
    se_rmst   = `se(rmean)`,
    lower_95  = rmst - 1.96 * se_rmst,
    upper_95  = rmst + 1.96 * se_rmst
  )

rmst_tbl
```

## Overall survival by subgroup

```{r os-subgroup}
model_surv_sub <- studyid_wise %>%
  mutate(
    time_os = as.numeric(os_censor - date_LTD1) / 365,
    status_os = if_else(os_censor_reason == "died", 1, 0),
    months_treat_time = as.numeric(difftime(sbrt_end, date_last_LDT, units = "days")) / 30.44,
    subgroup_flg = case_when(
      months_treat_time < 6  ~ 1,
      months_treat_time >= 6 ~ 2,
      TRUE ~ NA_real_
    ),
    subgroup = factor(subgroup_flg,
                      levels = c(1, 2),
                      labels = c("<6 months treatment", "≥6 months treatment"))
  )

model_surv_sub <- filter(model_surv_sub, !is.na(subgroup))

fit_os_subgroup <- survfit(Surv(time_os, status_os) ~ subgroup, data = model_surv_sub)

plot_os_subgroup <- ggsurvplot(
  fit_os_subgroup,
  conf.int      = FALSE,
  risk.table    = TRUE,
  pval          = TRUE,
  xlab          = "Years From First Recurrence",
  ylab          = "Overall Survival (%)",
  title         = "Kaplan–Meier Plot for Overall Survival by Subgroup",
  legend.title  = "",
  xlim          = c(0, 4),
  break.time.by = 1,
  ggtheme       = theme_minimal()
)

plot_os_subgroup$plot <- plot_os_subgroup$plot +
  scale_x_continuous(limits = c(0, 4), breaks = 0:4) +
  theme(
    panel.border     = element_rect(color = "black", fill = NA, linewidth = 0.8),
    panel.background = element_blank(),
    panel.grid       = element_blank()
  )

print(plot_os_subgroup)

log_rank <- survdiff(Surv(time_os, status_os) ~ subgroup, data = model_surv_sub)
pval <- 1 - pchisq(log_rank$chisq, df = length(log_rank$n) - 1)
cat("Log-rank p-value:", signif(pval, 4), "\n")
```

## Subgroup survival probabilities and RMST

```{r os-subgroup-probs-rmst}
times_vec <- 1:4

fit_overall <- survfit(Surv(time_os, status_os) ~ 1, data = model_surv_sub)
sp_all <- summary(fit_overall, times = times_vec)

overall_tbl <- tibble(
  group     = "All patients",
  time_year = sp_all$time,
  surv_prob = sp_all$surv,
  lower_95  = sp_all$lower,
  upper_95  = sp_all$upper
)

get_probs <- function(dat, label, times = times_vec, conf.type = "log-log") {
  sf  <- survfit(Surv(time_os, status_os) ~ 1, data = dat, conf.type = conf.type)
  ss  <- summary(sf, times = times)
  tibble(
    group     = label,
    time_year = ss$time,
    surv_prob = ss$surv,
    lower_95  = ss$lower,
    upper_95  = ss$upper
  )
}

by_group_tbl <- bind_rows(
  get_probs(filter(model_surv_sub, subgroup == "<6 months treatment"), "<6 months treatment"),
  get_probs(filter(model_surv_sub, subgroup == "≥6 months treatment"), "≥6 months treatment")
)

surv_probs_tbl <- bind_rows(overall_tbl, by_group_tbl) %>%
  mutate(across(c(surv_prob, lower_95, upper_95), ~ round(.x, 3)))

surv_probs_tbl

cutoff_tau_before_heavy_censoring <- function(fit, by = 0.25, risk_threshold = 0.10) {
  s_all <- summary(fit)
  t_max <- min(tapply(s_all$time, s_all$strata, max, na.rm = TRUE), na.rm = TRUE)
  times <- seq(0, t_max, by = by)

  s <- summary(fit, times = times)

  df <- tibble(
    time     = s$time,
    strata   = s$strata,
    n_risk   = s$n.risk
  )

  n0 <- df %>%
    group_by(strata) %>%
    summarise(N0 = max(n_risk, na.rm = TRUE), .groups = "drop")

  df2 <- df %>%
    inner_join(n0, by = "strata") %>%
    mutate(frac_at_risk = n_risk / N0)

  limiting <- df2 %>%
    group_by(time) %>%
    summarise(min_frac_at_risk = min(frac_at_risk, na.rm = TRUE), .groups = "drop")

  tau_suggested <- limiting %>%
    filter(min_frac_at_risk >= risk_threshold) %>%
    summarise(tau = max(time, na.rm = TRUE)) %>%
    pull(tau)

  list(
    tau_suggested       = tau_suggested,
    limiting_support    = limiting,
    per_time_support    = df2,
    t_max_shortest      = t_max,
    risk_threshold_used = risk_threshold,
    grid_by             = by
  )
}

out_tau <- cutoff_tau_before_heavy_censoring(fit_os_subgroup, by = 0.25, risk_threshold = 0.10)
tau_os_sub <- out_tau$tau_suggested
tau_os_sub

median_tbl <- summary(fit_os_subgroup)$table[, "median", drop = FALSE] %>%
  as.data.frame() %>%
  rownames_to_column("group") %>%
  rename(median_time = median)

rmst_tbl <- summary(fit_os_subgroup, rmean = tau_os_sub)$table %>%
  as.data.frame() %>%
  rownames_to_column("group") %>%
  transmute(
    group,
    n        = records,
    events   = events,
    tau      = tau_os_sub,
    rmst     = rmean,
    se_rmst  = `se(rmean)`,
    lower_95 = rmean - 1.96 * se_rmst,
    upper_95 = rmean + 1.96 * se_rmst
  ) %>%
  mutate(across(c(tau, rmst, se_rmst, lower_95, upper_95), ~ round(., 2)))

final_tbl <- median_tbl %>%
  left_join(rmst_tbl, by = "group")

final_tbl
```

# Cox regression analyses


```{r more-cox-univariable}
model_set1 <- model_set1 %>%
  mutate(
    hcc_size_gt = if_else(hcc_size_sbrt > 3, 1L, 0L)
  )

cox_hccsize <- coxph(
  Surv(time_lc_yr, resp_fail) ~ hcc_size_gt,
  data = model_set1
)
summary(cox_hccsize)

cox_trt_os <- coxph(
  Surv(time_os, status_os) ~ Desc_Salv_Consol,
  data = model_overall
)
summary(cox_trt_os)

model_overall <- model_overall %>%
  mutate(age_gt60 = if_else(age_first_SBRT > 60, 1L, 0L))

cox_age_gt60 <- coxph(
  Surv(time_os, status_os) ~ age_gt60,
  data = model_overall
)
summary(cox_age_gt60)

model_overall <- model_overall %>% 
  mutate(
    Milan_criteria = relevel(factor(Milan_criteria), ref = "No")
  )

cox_milan <- coxph(
  Surv(time_os, status_os) ~ Milan_criteria,
  data = model_overall
)
summary(cox_milan)

cox_listed <- coxph(
  Surv(time_os, status_os) ~ listed,
  data = model_overall
)
summary(cox_listed)

cox_transplant <- coxph(
  Surv(time_os, status_os) ~ transplant,
  data = model_overall
)
summary(cox_transplant)

cox_pvtt <- coxph(
  Surv(time_os, status_os) ~ pvtt,
  data = model_overall
)
summary(cox_pvtt)

model_overall <- model_overall %>%
  mutate(
    afp_prior_sbrt_cat = if_else(afp_prior_sbrt >= 10, 1L, 0L)
  )

table(model_overall$afp_prior_sbrt_cat, useNA = "ifany")

cox_afp <- coxph(
  Surv(time_os, status_os) ~ afp_prior_sbrt_cat,
  data = model_overall
)
summary(cox_afp)

model_overall <- model_overall %>%
  mutate(
    cp_prior_group = relevel(factor(cp_prior_group), ref = "B/C")
  )

cox_cp <- coxph(
  Surv(time_os, status_os) ~ cp_prior_group,
  data = model_overall
)
summary(cox_cp)

model_overall <- model_overall %>%
  mutate(
    hcc_size_gt = if_else(hcc_size_sbrt > 3, 1L, 0L)
  )

cox_hccsize <- coxph(
  Surv(time_os, status_os) ~ hcc_size_gt,
  data = model_overall
)
summary(cox_hccsize)

model_overall <- model_overall %>%
  mutate(
    eqd2_gt80 = if_else(eqd2 > 80, 1L, 0L)
  )

cox_eqd2 <- coxph(
  Surv(time_os, status_os) ~ eqd2_gt80,
  data = model_overall
)
summary(cox_eqd2)
```

## Multivariable Cox model

```{r cox-multivariable}
model_overall <- model_overall %>%
  mutate(
    Desc_Salv_Consol = relevel(factor(Desc_Salv_Consol), ref = "Salvage"),
    hcc_size_gt      = factor(if_else(hcc_size_sbrt > 3, 1L, 0L)),
    age_gt60         = factor(if_else(age_first_SBRT > 60, 1L, 0L)),
    Milan_criteria   = relevel(factor(Milan_criteria), ref = "No"),
    listed           = factor(listed),
    transplant       = factor(transplant),
    pvtt             = factor(pvtt),
    cp_prior_group   = relevel(factor(cp_prior_group), ref = "B/C"),
    afp_prior_sbrt_cat = factor(if_else(afp_prior_sbrt >= 10, 1L, 0L)),
    eqd2_gt80        = factor(if_else(eqd2 > 80, 1L, 0L))
  )

forced_vars <- c("Desc_Salv_Consol", "hcc_size_gt")
candidates  <- c("age_gt60","Milan_criteria","listed","transplant","pvtt",
                 "afp_prior_sbrt_cat","cp_prior_group","eqd2_gt80")

dat <- model_overall %>%
  select(time_os, status_os, all_of(forced_vars), all_of(candidates)) %>%
  na.omit()

base_f <- as.formula(
  paste("Surv(time_os, status_os) ~", paste(forced_vars, collapse = " + "))
)

full_f <- as.formula(
  paste("Surv(time_os, status_os) ~",
        paste(c(forced_vars, candidates), collapse = " + "))
)

fit_base <- coxph(base_f, data = dat, x = TRUE, y = TRUE)

fit_forward_aic <- step(
  fit_base,
  scope = list(lower = base_f, upper = full_f),
  direction = "forward",
  trace = 1
)

summary(fit_forward_aic)

hr_tbl <- data.frame(
  Variable = rownames(coef(summary(fit_forward_aic))),
  HR       = exp(coef(fit_forward_aic)),
  LCL      = exp(confint(fit_forward_aic)[,1]),
  UCL      = exp(confint(fit_forward_aic)[,2]),
  p_value  = coef(summary(fit_forward_aic))[,"Pr(>|z|)"],
  row.names = NULL
)

hr_tbl
cox.zph(fit_forward_aic)
```

## Manual model building

```{r cox-manual}
dat <- model_overall %>%
  mutate(
    Desc_Salv_Consol = relevel(factor(Desc_Salv_Consol), ref = "Salvage"),
    hcc_size_gt      = factor(if_else(hcc_size_sbrt > 3, 1L, 0L))
  )

fit_start <- coxph(Surv(time_os, status_os) ~ Desc_Salv_Consol + hcc_size_gt,
                   data = dat)
summary(fit_start)
AIC(fit_start)

fit_2 <- coxph(Surv(time_os, status_os) ~ Desc_Salv_Consol + hcc_size_gt + age_gt60,
               data = dat)
summary(fit_2)
AIC(fit_2)

fit_3 <- coxph(Surv(time_os, status_os) ~ Desc_Salv_Consol + hcc_size_gt + pvtt,
               data = dat)
summary(fit_3)
AIC(fit_3)

fit_4 <- coxph(Surv(time_os, status_os) ~ Desc_Salv_Consol + hcc_size_gt + pvtt + afp_prior_sbrt_cat,
               data = dat)
summary(fit_4)
AIC(fit_4)

fit_5 <- coxph(Surv(time_os, status_os) ~ Desc_Salv_Consol + hcc_size_gt + pvtt + afp_prior_sbrt_cat + Milan_criteria,
               data = dat)
summary(fit_5)
AIC(fit_5)

fit_6 <- coxph(Surv(time_os, status_os) ~ Desc_Salv_Consol + hcc_size_gt + pvtt + afp_prior_sbrt_cat + listed,
               data = dat)
summary(fit_6)
AIC(fit_6)

fit_7 <- coxph(Surv(time_os, status_os) ~ Desc_Salv_Consol + hcc_size_gt + pvtt + afp_prior_sbrt_cat + listed + transplant,
               data = dat)
summary(fit_7)
AIC(fit_7)

fit_8 <- coxph(Surv(time_os, status_os) ~ Desc_Salv_Consol + hcc_size_gt + pvtt + afp_prior_sbrt_cat + listed + transplant + cp_prior_group,
               data = dat)
summary(fit_8)
AIC(fit_8)

fit_9 <- coxph(Surv(time_os, status_os) ~ Desc_Salv_Consol + hcc_size_gt + pvtt + afp_prior_sbrt_cat + listed + transplant + cp_prior_group + eqd2_gt80,
               data = dat)
summary(fit_9)
AIC(fit_9)

final_model_os <- coxph(Surv(time_os, status_os) ~ Desc_Salv_Consol + hcc_size_gt + pvtt + afp_prior_sbrt_cat + listed + transplant + cp_prior_group,
                        data = dat)
summary(final_model_os)
AIC(final_model_os)
```

# Additional analyses

## Recurrent vs residual tumors

```{r recurrent-vs-residual}
model_lc <- model_set1 %>%
  filter(residual_reccurence %in% c("Residual", "Recurrence", "Not included"))

table9_summary <- model_lc %>%
  group_by(residual_reccurence) %>%
  summarise(N = n(), .groups = "drop") %>%
  mutate(Percent = round(100 * N / sum(N), 1))

table9_summary

model_km <- model_set1 %>%
  mutate(
    group_recur = if_else(residual_reccurence == "Recurrence",
                          "Recurrent tumor followed by salvage SBRT",
                          "CMT including residual tumors")
  )

fit_km <- survfit(Surv(time_lc_yr, resp_fail) ~ group_recur, data = model_km)

lc_summary <- summary(fit_km, times = c(1:4))
km_table <- data.frame(
  Group   = rep(levels(factor(model_km$group_recur)), each = length(lc_summary$time) / 2),
  Time    = lc_summary$time,
  LC_rate = round(lc_summary$surv * 100, 1)
)

print(km_table)

plot_recurrent_group <- ggsurvplot(
  fit_km,
  data = model_km,
  risk.table = TRUE,
  pval = sprintf("p = %.2f", surv_pvalue(fit_km, data = model_km)$pval),
  conf.int = FALSE,
  xlab = "Years from First Recurrence",
  ylab = "Local Control (%)",
  title = "Kaplan–Meier Plot of Local Control by Treatment Type in Recurrent vs. Residual Tumors",
  legend.title = "",
  legend.labs = c("CMT including residual tumors", "Recurrent tumor followed by salvage SBRT"),
  break.time.by = 1,
  xlim = c(0, 4),
  ggtheme = theme_minimal()
)

plot_recurrent_group$plot <- plot_recurrent_group$plot +
  scale_x_continuous(limits = c(0, 4), breaks = 0:4) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.8),
    panel.background = element_blank(),
    panel.grid = element_blank()
  )

print(plot_recurrent_group)
```

## Recurrent vs residual probability table and RMST

```{r recurrent-vs-residual-rmst}
fit_km <- survfit(
  Surv(time_lc_yr, resp_fail) ~ group_recur,
  data = model_km,
  conf.type = "log-log"
)

lc_summary <- summary(fit_km, times = c(1, 2, 3, 4))

lc_df <- data.frame(
  Group    = lc_summary$strata,
  Time     = lc_summary$time,
  N_Risk   = lc_summary$n.risk,
  N_Event  = lc_summary$n.event,
  Survival = round(lc_summary$surv  * 100, 1),
  LCL      = round(lc_summary$lower * 100, 1),
  UCL      = round(lc_summary$upper * 100, 1),
  row.names = NULL
)

lc_df$Group <- sub("^.*=", "", lc_df$Group)
lc_df

summary(fit_km)$table

max_by_arm <- with(model_km,
                   tapply(time_lc_yr, group_recur, max, na.rm = TRUE))
max_by_arm

tau <- 4.5

rmst_tbl <- summary(fit_km, rmean = tau)$table |>
  as.data.frame(check.names = FALSE) |>
  tibble::rownames_to_column("group") |>
  dplyr::transmute(
    group     = sub("^.*=", "", group),
    tau       = tau,
    rmst      = rmean,
    se_rmst   = `se(rmean)`,
    lower_95  = rmst - 1.96 * se_rmst,
    upper_95  = rmst + 1.96 * se_rmst
  )
rmst_tbl
```

## Tau selection before heavy censoring for recurrent vs residual

```{r recurrent-tau}
cutoff_tau_before_heavy_censoring <- function(fit, by = 0.25, risk_threshold = 0.10) {
  s_all <- summary(fit)
  t_max <- min(tapply(s_all$time, s_all$strata, max, na.rm = TRUE), na.rm = TRUE)
  times <- seq(0, t_max, by = by)

  s <- summary(fit, times = times)

  df <- tibble(
    time   = s$time,
    strata = s$strata,
    n_risk = s$n.risk
  )

  n0 <- df %>%
    group_by(strata) %>%
    summarise(N0 = max(n_risk, na.rm = TRUE), .groups = "drop")

  df2 <- df %>%
    inner_join(n0, by = "strata") %>%
    mutate(frac_at_risk = n_risk / N0)

  limiting <- df2 %>%
    group_by(time) %>%
    summarise(min_frac_at_risk = min(frac_at_risk, na.rm = TRUE), .groups = "drop")

  tau_suggested <- limiting %>%
    filter(min_frac_at_risk >= risk_threshold) %>%
    summarise(tau = max(time, na.rm = TRUE)) %>%
    pull(tau)

  list(
    tau_suggested       = tau_suggested,
    limiting_support    = limiting,
    per_time_support    = df2,
    t_max_shortest      = t_max,
    risk_threshold_used = risk_threshold,
    grid_by             = by
  )
}

out_tau_lc <- cutoff_tau_before_heavy_censoring(fit_km, by = 0.25, risk_threshold = 0.10)
tau_lc <- out_tau_lc$tau_suggested
tau_lc

rmst_lc_tbl <- summary(fit_km, rmean = tau_lc)$table %>%
  as.data.frame() %>%
  rownames_to_column("group") %>%
  transmute(
    group,
    n        = records,
    events   = events,
    tau      = tau_lc,
    rmst     = rmean,
    se_rmst  = `se(rmean)`,
    lower_95 = rmean - 1.96 * se_rmst,
    upper_95 = rmean + 1.96 * se_rmst
  ) %>%
  mutate(across(c(tau, rmst, se_rmst, lower_95, upper_95), ~ round(., 2)))

rmst_lc_tbl
```

## CMT diagnostic imaging subgroup

```{r cmt-diagnostic-subgroup}
model_set1 <- model_set1 %>%
  mutate(
    cmt_diag_test_bin = case_when(
      cmt_diag_test == "No"  ~ 0,
      cmt_diag_test == "Yes" ~ 1,
      cmt_diag_test == "Not included" ~ NA_real_
    )
  )

table(model_set1$cmt_diag_test, model_set1$cmt_diag_test_bin, useNA = "ifany")

cmt_subset <- model_set1 %>%
  filter(Desc_Salv_Consol == "Consolidate/CMV") %>%
  mutate(
    cmt_diag_test_bin = dplyr::case_when(
      cmt_diag_test == "No"  ~ 0,
      cmt_diag_test == "Yes" ~ 1,
      TRUE ~ NA_real_
    ),
    diag_group = if_else(cmt_diag_test_bin == 1,
                         "Diagnostic imaging between tx",
                         "No diagnostic imaging between tx")
  ) %>%
  filter(!is.na(cmt_diag_test_bin))

table(cmt_subset$diag_group, useNA = "ifany")
```

## CMT diagnostic imaging KM and RMST

```{r cmt-diagnostic-km}
fit_km_cmt <- survfit(
  Surv(time_lc_yr, resp_fail) ~ diag_group,
  data = cmt_subset,
  conf.type = "log-log"
)

lc_summary <- summary(fit_km_cmt, times = c(1, 2, 3, 4))

lc_df <- data.frame(
  Group    = sub("^.*=", "", lc_summary$strata),
  Time     = lc_summary$time,
  N_Risk   = lc_summary$n.risk,
  N_Event  = lc_summary$n.event,
  Survival = round(lc_summary$surv  * 100, 1),
  LCL      = round(lc_summary$lower * 100, 1),
  UCL      = round(lc_summary$upper * 100, 1),
  row.names = NULL
)
lc_df

cmt_trunc <- cmt_subset %>%
  mutate(
    time4   = pmin(time_lc_yr, 4),
    status4 = ifelse(resp_fail == 1 & time_lc_yr <= 4, 1, 0)
  )

fit_km_cmt_trim <- survfit(Surv(time4, status4) ~ diag_group,
                           data = cmt_trunc, conf.type = "log-log")

km <- ggsurvplot(
  fit_km_cmt_trim,
  data = cmt_trunc,
  conf.int   = FALSE,
  pval       = TRUE,
  risk.table = TRUE,
  xlim       = c(0, 4),
  title = "Kaplan–Meier Plot of Local Control in CMT Tumors With and Without Diagnostic Imaging Between Treatments",
  break.time.by = 1,
  ggtheme    = theme_minimal(),
  xlab       = "Years from First Recurrence",
  ylab       = "Local Control (%)",
  legend.title = "",
  legend.labs  = c("Diagnostic imaging between tx", "No diagnostic imaging between tx")
)

km$plot <- km$plot +
  scale_x_continuous(limits = c(0, 4), breaks = 0:4) +
  theme(
    panel.border     = element_rect(color = "black", fill = NA, linewidth = 0.8),
    panel.background = element_blank(),
    panel.grid       = element_blank()
  )

print(km)

tapply(cmt_subset$time_lc_yr, cmt_subset$diag_group, max)

tau <- min(
  tapply(cmt_subset$time_lc_yr, cmt_subset$diag_group, max)
)
tau

cmt_subset$diag_bin <- ifelse(
  cmt_subset$diag_group == "Diagnostic imaging between tx", 1, 0
)

rmst_tbl <- summary(fit_km_cmt, rmean = tau)$table |>
  as.data.frame(check.names = FALSE) |>
  tibble::rownames_to_column("group") |>
  dplyr::transmute(
    group     = sub("^.*=", "", group),
    tau       = tau,
    rmst      = rmean,
    se_rmst   = `se(rmean)`,
    lower_95  = rmst - 1.96 * se_rmst,
    upper_95  = rmst + 1.96 * se_rmst
  )
rmst_tbl
```

## Alternative tau selection for CMT subgroup

```{r cmt-tau}
km_df <- broom::tidy(fit_km_cmt) %>%
  group_by(strata) %>%
  mutate(risk_prop = n.risk / max(n.risk)) %>%
  summarise(
    tau_suggested = max(time[risk_prop >= 0.10]),
    .groups = "drop"
  )

tau_cmt <- min(km_df$tau_suggested)
tau_cmt

rmst_cmt_tbl <- summary(fit_km_cmt, rmean = tau_cmt)$table %>%
  as.data.frame() %>%
  tibble::rownames_to_column("group") %>%
  transmute(
    group,
    n        = records,
    events   = events,
    tau      = tau_cmt,
    rmst     = rmean,
    se_rmst  = `se(rmean)`,
    lower_95 = rmean - 1.96 * se_rmst,
    upper_95 = rmean + 1.96 * se_rmst
  ) %>%
  mutate(across(c(tau, rmst, se_rmst, lower_95, upper_95), ~ round(., 2)))

rmst_cmt_tbl
```

## Table 10 interval summary

```{r table-10}
model_set1 <- model_set1 %>%
  mutate(
    sbrt_end = if_else(
      unique_id == 77,
      as.Date("2018-10-19"),
      sbrt_end
    )
  )

model_set1 %>%
  filter(unique_id == 77) %>%
  select(unique_id, date_last_LDT, sbrt_end)

model_set1$ldt_to_sbrt_interval_months <- time_length(
  interval(model_set1$date_last_LDT, model_set1$sbrt_end),
  unit = "months"
)

group_summary <- model_set1 %>%
  group_by(Desc_Salv_Consol) %>%
  summarise(
    Median = round(median(ldt_to_sbrt_interval_months, na.rm = TRUE), 1),
    Q1 = round(quantile(ldt_to_sbrt_interval_months, 0.25, na.rm = TRUE), 1),
    Q3 = round(quantile(ldt_to_sbrt_interval_months, 0.75, na.rm = TRUE), 1),
    Range_Low = round(min(ldt_to_sbrt_interval_months, na.rm = TRUE), 1),
    Range_High = round(max(ldt_to_sbrt_interval_months, na.rm = TRUE), 1),
    N = sum(!is.na(ldt_to_sbrt_interval_months)),
    .groups = "drop"
  ) %>%
  mutate(
    IQR = paste0("(", Q1, "–", Q3, ")"),
    Range = paste0("(", Range_Low, "–", Range_High, ")"),
    p_value = NA
  ) %>%
  select(Group = Desc_Salv_Consol, Median, IQR, Range, N, p_value)

total_summary <- model_set1 %>%
  summarise(
    Median = round(median(ldt_to_sbrt_interval_months, na.rm = TRUE), 1),
    Q1 = round(quantile(ldt_to_sbrt_interval_months, 0.25, na.rm = TRUE), 1),
    Q3 = round(quantile(ldt_to_sbrt_interval_months, 0.75, na.rm = TRUE), 1),
    Range_Low = round(min(ldt_to_sbrt_interval_months, na.rm = TRUE), 1),
    Range_High = round(max(ldt_to_sbrt_interval_months, na.rm = TRUE), 1),
    N = sum(!is.na(ldt_to_sbrt_interval_months))
  ) %>%
  mutate(
    Group = "Total",
    IQR = paste0("(", Q1, "–", Q3, ")"),
    Range = paste0("(", Range_Low, "–", Range_High, ")"),
    p_value = round(wilcox.test(ldt_to_sbrt_interval_months ~ Desc_Salv_Consol, data = model_set1)$p.value, 6)
  ) %>%
  select(Group, Median, IQR, Range, N, p_value)

final_summary <- bind_rows(group_summary, total_summary)
final_summary
```
